---
title: "Car Sale Features Per Year since 2000"
format: html
editor: visual
execute: 
  warning: false
  message: false
  echo: false
---

```{r}
#| echo: false
#| include: false
library(tidyverse)
library(scales)
library(ggthemes)
car_data <- read.csv("data/car-dataset.csv", sep = ",")

car_data <- car_data %>%
  mutate(
    Year = as.factor(Year),
    Country = as.factor(Country),
    Car_Brand = as.factor(Car_Brand),
    Fuel_Type = as.factor(Fuel_Type),
    Units_Sold = as.numeric(Units_Sold),
    Average_Price = as.numeric(Average_Price),
    Safety_Rating = as.factor(Safety_Rating),
    Horsepower = as.numeric(Horsepower)
  )
```

```{r plot1}
car_data %>%
  group_by(Year) %>%
  summarise(Avg_Price = mean(Average_Price, na.rm = TRUE)) %>%
  ggplot(aes(x = as.numeric(as.character(Year)), y = Avg_Price)) +
  geom_line(color = "#2E86AB", size = 1.5) +
  geom_point(color = "#E74C3C", size = 3) +
  geom_smooth(method = "loess", se = FALSE, color = "darkgreen", linetype = "dashed") +
  labs(title = "Trend of Average Car Price Over the Years", x = "Year", y = "Average Price (USD)") +
  scale_y_continuous(labels = dollar) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    panel.grid.minor = element_blank()
  )
```

This graph shows a clear trend: the average price of cars has been steadily increasing every 5 years, from 
25.000 in 2000 to 37,500 in 2025—a 50% jump in just 25 years.

***Key Reasons for Rising Prices:***

**Technology & Features**
Modern cars now include advanced safety systems (automatic braking, lane assist), touchscreen infotainment, and hybrid/electric options—all adding to the cost.

**Inflation & Production Costs**

Materials (steel, aluminum, lithium for batteries) and labor costs have risen over time.
Stricter emissions and safety regulations require expensive engineering.

**Consumer Preferences**

Buyers want SUVs and trucks (which cost more than small cars).
Luxury brands and premium features are in higher demand.

**Supply Chain Issues**

Events like the pandemic and chip shortages (2020–2022) disrupted production, pushing prices higher.

***What This Means for You:***

**New cars** are becoming a bigger investment. Many buyers now opt for longer loans (6–7 years) or used vehicles.

**Electric cars,** while pricier upfront, may save money on fuel and maintenance long-term.

**Leasing or subscriptions** are growing as alternatives to buying.



```{r plot2}
# Get most sold fuel type per country and year
most_sold_fuel_country_year <- car_data %>%
  group_by(Year, Country, Fuel_Type) %>%
  summarise(Total_Units_Sold = sum(Units_Sold, na.rm = TRUE), .groups = "drop") %>%
  group_by(Year, Country) %>%
  slice_max(Total_Units_Sold, n = 1)  # only most sold fuel type

# Plot
ggplot(most_sold_fuel_country_year, aes(x = Year, y = fct_rev(Country), fill = Fuel_Type)) +
  geom_tile(color = "white", width = 0.9, height = 0.9) +
  scale_fill_brewer(palette = "Set2", name = "Fuel Type") +
  labs(
    title = "Most Sold Fuel Type by Country and Year",
    x = "Year", y = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 11),
    axis.text.y = element_text(size = 11),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )
```

This graph shows which **fuel types** (Diesel, Electric, Hybrid, Petrol) were the **most sold** in different countries **(Canada, France, Germany, Italy, Japan, Russia, UK, USA)** from **2000 to 2024.**

**Key Trends to Notice:**
**Early 2000s (2000–2010):**

**Petrol (gasoline) and Diesel** were dominant in most countries.

**Diesel** was especially popular in Europe (France, Germany, Italy, UK).

**2010s (2010–2020):**

**Hybrid cars** (like Toyota Prius) grew in popularity, especially in **Japan and the USA.**

**Electric cars (EVs)** started appearing but were still rare.

**2020s (2020–2024):**

**Electric cars (EVs) surged** due to better technology, government incentives, and environmental awareness.

**Diesel sales dropped** in Europe due to pollution concerns.

**Hybrids remained strong** as a transition option.

**What This Means:**
**Petrol (Gasoline)** was the default choice, but **electric and hybrid cars are taking over.**

**Europe** moved from **Diesel → Electric.**

**USA & Japan** shifted from **Petrol → Hybrid → Electric.**

**Russia** still relies more on **Petrol/Diesel** (slower EV adoption).

**Future Prediction:**
Electric cars will likely become the #1 fuel type in most countries by **2030.**


```{r plot3}
# Load libraries
library(tidyverse)
library(readr)
library(scales)

# Read the dataset
car_data <- read_csv("data/car-dataset.csv")

# Define custom horsepower bins: start at 50, step by 75
hp_min <- 50
hp_max <- max(car_data$Horsepower, na.rm = TRUE)
hp_breaks <- seq(hp_min, ceiling(hp_max + 75), by = 75)

# Create labeled HP ranges
car_data <- car_data %>%
  mutate(HP_Range = cut(Horsepower, breaks = hp_breaks, include.lowest = TRUE))

# Summarize Units Sold by HP Range and Safety Rating
grouped_data <- car_data %>%
  group_by(HP_Range, Safety_Rating) %>%
  summarise(Total_Units = sum(Units_Sold, na.rm = TRUE), .groups = "drop")

# Create the faceted plot
ggplot(grouped_data, aes(x = HP_Range, y = Total_Units, fill = Safety_Rating)) +
  geom_bar(stat = "identity", alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ Safety_Rating, ncol = 2) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = "Units Sold by Horsepower Range",
    x = "Horsepower Range",
    y = "Total Units Sold"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 14, face = "bold")
  )
```

This graph shows *how many cars were sold* based on their *horsepower (HP)* and *safety ratings (A, B, C, D), where:*

**A = Best safety | B = Good | C = Average | D = Poor**

*HP ranges* span from *100 HP to 400+ HP*

***Key Trends:***

**Most Popular Horsepower Range:**

The majority of cars sold fall in the *150–250 HP* range.

This balances *power, fuel efficiency, and affordability* for everyday drivers.

**Safety vs. Performance:**

*Highest sales* are in *A/B-rated* cars with *moderate HP (150–250)*—safety matters most to average buyers.

*High-HP cars (300+ HP)* sell in smaller numbers and often have *lower safety ratings (C/D)*—common in sports/luxury vehicles.

*Low-HP cars (<150 HP)* may prioritize fuel economy but vary in safety.

**Market Preferences:**

*Family/commuter cars:* High safety (A/B) + mid-range HP.

*Performance cars:* High HP (300+) but often trade safety for speed.

*Budget cars:* Lower HP + mixed safety ratings.

**What This Means for Buyers:**
✔ *Need a safe, practical car?* Look for A/B ratings with 150–250 HP.
✔ *Want speed?* High-HP cars exist but may cost more and compromise safety.
✔ *Shopping on a budget?* Lower HP options are available, but check safety ratings carefully.
```{r plot4}
# Summarize: Get most sold car color per year
most_sold_colors <- car_data %>%
  group_by(Year, Car_Color) %>%
  summarise(Units_Sold = sum(Units_Sold, na.rm = TRUE), .groups = "drop") %>%
  group_by(Year) %>%
  slice_max(Units_Sold, n = 1)

# Plot: Bar chart with legend on right
ggplot(most_sold_colors, aes(x = Year, y = Units_Sold, fill = Car_Color)) +
  geom_col(width = 0.7) +  # show.legend = TRUE by default
  labs(
    title = "Most Popular Car Color Each Year (by Units Sold)",
    x = "Year",
    y = "Units Sold",
    fill = "Car Color"
  ) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.1))) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 75, hjust = 1),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.position = "right",                 # move legend to right of plot
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

```

A bar chart displaying the top-selling car color for each year based on total units sold.

