---
title: "Home Page"
---
Welcome to My Car Sales Data Analysis Project
This website presents a comprehensive analysis of car sales data from the year 2000 onward. Inside, you'll find interactive visualizations and clear summaries highlighting key trends in the automotive market—such as the most popular car colors, top-selling vehicle features, and year-over-year changes in consumer preferences. Explore the Data Analysis section for detailed graphs and insights, check out the Sources page for data provenance, and visit the About section to learn more about the purpose and scope of this project.



```{r setup} 
#| include: false
library(ggplot2)
library(dplyr)
```

```{r}
#| echo: false
car_data <- read.csv("data/car-dataset.csv")
```

```{r plot}
#| echo: false
#| warning: false
#| fig-align: center
#| fig-width: 8
#| fig-height: 6

car_data %>%
  group_by(Year) %>%
  summarise(Total_Sold = sum(Units_Sold, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = Total_Sold, group = 1)) +
  geom_line(color = "darkblue", size = 1.2) +
  geom_point(color = "red", size = 3) +
  labs(title = "Total Units Sold Over Time",
       x = "Year",
       y = "Units Sold") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"))
```


